---
title: "Geog533 Lab 10"
author: "Tyler Sherven"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
if(!require(UScensus2010)) {install.packages("UScensus2010")}
library(UScensus2010)
if(!require(UScensus2010county)) {install.county("osx")}
library(UScensus2010county)
library(tmap)

data("new_york.county10")
nys <- new_york.county10
plot(nys)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
install.packages("tmap")
library(tmap)
qtm(nys)
```


### Question 1(c)	
How many counties in New York State?
```{r}
conys <- coordinates(nys)
datany <- nys@data
length(datany$NAME10)
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
datany[1,4]
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
nyspop <- datany$P0010001

mean(nyspop)
sum(nyspop)
median(nyspop)
max(nyspop)
min(nyspop)

library(moments)
skewness(nyspop)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(datany$P0010001)
boxplot(datany$P0010001)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r}
if(!require(UScensus2010tract)) {install.tract("osx")}
library(UScensus2010tract)
```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
data("new_york.tract10")
nys2 <- new_york.tract10
plot(nys2)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
nysco2 <- coordinates(nys2)
nyspop2 <- nys2@data
sum(nyspop2$P0010001)
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
bct <- nys2[nys2$county=="007",]
plot(bct)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
bpop <- nys2[nys2$county=="007",]
sum(bpop$P0010001)
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(bpop$P0010001)
sum(bpop$P0010001)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
select <- bpop[,1:5]
select$ratio <- bpop$P0010001/{sum(bpop$P0010001)}
select
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r}
if(!require(ncdf4)) install.packages("ncdf4")
library(raster)
library(ncdf4)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
rasbn <- raster("NDVI.nc")
plot(rasbn)
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(rasbn)
ncol(rasbn)
ncell(rasbn)
nbands(rasbn)
res(rasbn)
extent(rasbn)
bbox(rasbn)
projection(rasbn)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
rasbnmin <- min(rasbn)
rasbmax <- max(rasbn)
writeRaster(rasbnmin,filename="rasbnmin.tif",overwrite=TRUE)
writeRaster(rasbmax,filename="rasbmax.tif",overwrite=TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
plot(rasbn,13:24)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r}
#hist(rasbn,13:24)
#Error in .local
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(rasbn, 7)

rasvalues <- rasbn[50,40]
plot(as.vector(rasvalues), type="b" )
```

