---
title: "Which Economic and Ethnic/Race Factors are Most Impactful on Median Property Values Broome County Census Tracts?"
author: "Tyler Sherven"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---


## 1 CSV & Packages
Csv file containing 2011-15 Broome County Census Tract ACS data for economic factors: Labor Force Participation Rate 16+ and Median Household Income, and race/ethnic census tract distribution of white, black, asian, native american, and latino populations. Install necessary packages.
```{r}
library(readr)
data1 <- read_csv("~/Documents/533FinalProjectData.csv")
data1

library(knitr)
library(car)
library(MASS)
```


## 2 Setting up variables for cleaner coding
```{r}
medpropvalue <- data1$MedPropValue
labor <- data1$`LaborForceParticipationRate16+`
medhouseinc <- data1$HouseholdMedInc
black <- data1$Black
white <- data1$White
asian <- data1$Asian
native <- data1$Native
latino <- data1$`Hispanic/Latino`
```


## 3 Linear Model Function and Summary including all variables
```{r}
a <- lm(medpropvalue~medhouseinc+labor+black+native+latino+white+asian)
summary(a)
##Economic factors of Labor Force Participation Rate 16+ and Household Median Income are most significant with 3 stars, while Asian Population is also significant with 3 stars.
```


## 4 Improved linear model function and summary of all variables
```{r}
b <- lm(medpropvalue~labor+medhouseinc+asian)
summary(b)
##Improved linear model results in Asian population ranking as the most significant variable with 3 stars, median household income ranking 2nd with 3 stars, and labor ranking last with no stars.
```


## 5 Linear Model and Summary including Race/Ethnic Variables
```{r}
c <- lm(medpropvalue~white+black+native+asian+latino)
summary(c)
##Asian, White, and Black populations rate as significant variables.
```


## 6 Improved Linear model and summary of Race/Ethnic Variables
```{r}
d <- lm(medpropvalue~white+black+asian)
summary(d)
##Asian Population is most significant in determining Broome Census Tract Property Values, followed by Black and then White Populations in signifiance with 2 stars.
```

## 7 Testing Null Hypothesis of each variable
Null Hypothesis: there is a relationship between Median Property Value and Variable. Alternative Hypothesis: there is no relationship between Median Property Value and Variable. P.Value less than .05, we accept null hypothesis.
```{r}
cor.test(black, medpropvalue, method = "pearson") #Null Hypothesis
cor.test(white, medpropvalue, method = "pearson") #Alternative Hypothesis
cor.test(asian, medpropvalue, method = "pearson") #Null Hypothesis
cor.test(labor, medpropvalue, method = "pearson") #Alternative Hypothesis
cor.test(medhouseinc, medpropvalue, method = "pearson") #Alternative Hypothesis
```


## 8 Data Visualization
Scatterplots to visualize the relationship between median property value and each significant variable from the original and race/ethnic linear models
```{r}
plot(medpropvalue, medhouseinc)
plot(medpropvalue, black)
plot(medpropvalue, white)
plot(medpropvalue, asian)
plot(medpropvalue, labor)
```

## 9 Conclusion
Variables that both demonstrate linear model significance and fulfill null hypothesis, ranking as most impactful on Broome County Median Property Value: 1. Asian  Population, 2. Black Population.