---
title: "Geog533 Lab 8"
author: "Tyler Sherven"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
resp <- 24

regsumsquare <- 2.3 - 1.7

totaldf <- resp-1
resdf <- resp-2

resmeansquare <- 1.7/resdf
resmeansquare1 <- round(resmeansquare, 3)

f <- round(regsumsquare/resmeansquare1)
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression| r regsumsquare  |          1         |r regssumsquare |  r F            |
| Residual |       1.7       |    r resdf         |r resmeansquare1|                 |
| Total    |       2.3       |    r totaldf       |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
.46 + (.10 * 50)
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

###Solution 1(c)
With each dollar an individual earns, they will embark on about .19 shopping trips each week


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
cr <- qf(0.95, 1, 22)
if(f < cr){
print("We accept the null hypothesis")
} else {
print("We reject the null hypothesis")}

### the regression coeffiecient is significantly different different than 0, since the F value is larger than the critical F value
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
r <- sqrt(regsumsquare/2.3)
r
```




## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
s <- c(36, 78, 11, 45)
e <- c(400, 800, 200, 675)

m <- lm(s~e)
summary(m)

###slope = .09665
###intercept = -7.63721
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
summary(m)

###Residual Standard Error = 11.25
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
crf <- qf(.95, 1, 2)
crf

###The critical F at 18.51282 is larger than the regression coefficient of 16.173. Therefore we can't reject the null hypothesis
```


### Question 2(d)
Find the value of $r^2$.
```{r}
d <- cor.test(e, s, method = "pearson")
d2 <- d$estimate^2
d2
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
edf <- data.frame(s, e)

edf$predicted <- m$fitted.values
edf$residuals <- m$residuals
edf

library(knitr)
kable(edf)

```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
anova(m)
2047 + 253.23
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|    2047.77      |       1            |   2047.77      |    16.173      |
| Residual |   253.23        |       2            |      126.61    |                 |
| Total    |    2300.23      |       3            |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(e,s)
abline(m,col = "green")
segments(e, fitted(m), e, s, col="red", lty = "dashed")
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"

a <- read.csv(url)
ms <- lm(a$SalePrice~a$LotSize)

summary(ms)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"

uka <- read.csv(url,header = TRUE)
ukb <- lm(uka$price~uka$bedrooms)

summary(ukb)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
b4 <- lm(uka$price~uka$bathrooms)
summary(b4)
```
The number of bathrooms has less of an impact on price than the number of bedrooms. These results lead us to reject the null hypothesis.

## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
library(leaflet)
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-78.87, lat=42.9659, popup="home")
m  # Print the map
```

